Transparency of COVID-19-related research in medical journals

### 2022-04-12 manuscript_covid_transparency.qmd
discussion

Adherence to transparency practices increased over time in COVID-19 related medical literature. Adherence to reporting COI disclosure was high throughout study period. Code sharing and protocol registration were rare, but improved little over study period.

@@

@@

>>>>

Adherence to transparency practices increased over time in COVID-19 related medical literature during 2020-2021. Adherence to reporting COI disclosure was high throughout study period, in addition most of articles had funding disclosure. Data sharing, code sharing and protocol registration were rare, but improved little over study period. Conflicts of interest disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparency practices according to received citations and JIF were small.

@@

In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodology and scale of our research. However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the begininning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\]. In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\]. Lacking preregisteration, sharing code and conflicts of interest have go attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research.

@@

Transparency is crucial to ensure the credibility of science and enable its assessment. It has been argued that open science, including transparency practices such as the ones we investigated here, have helped to fight pandemic globally [@besançon2021; @vuong2022]. However, your findings showed suboptimal adherence to data and code sharing and pre-registeration in COVID-19 related research, whilst majority of researchers had adhered to disclose funding and conflicts of interests related to their COVID-19 related medical research. Stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
